---
title: 'Rush Trial: data exploration 2017'
output:  html_notebook
#  html_document:
#    keep_md: yes

---

Packages
```{r}
library(tidyverse)
library(skimr)
library(funModeling)
```

Import data
```{r}
grp_calc_w <- read.csv("../data/prepped/taxon_grp_calc_w.csv", header = TRUE)
```

```{r}
#drop unwanted columns
data_taxa <- grp_calc_w %>% 
  select(-c(X, uid)) 

#make categories factors
data_taxa <- data_taxa %>% 
  mutate_at(c("year", "treat_plot"), as.factor)

skim(data_taxa)  
funModeling::plot_num(data_taxa) #plot numeric variables
funModeling::profiling_num(data_taxa) %>% pander() #instead of pander use kable()
funModeling::correlation_table(data = data_taxa, target = "treat_plot") #correlation coefficient (pearsons R)
```


filter by year and type
```{r}
# data_taxa <- filter(data_taxa, year == "2015", type == "meadow")
# 
# #drop unused factors
# data_taxa <- droplevels(data_taxa)
# 
# skim(data_taxa)

```

# Analysis

##Three-way ANOVA with replication

### Baseline data (2015) 
```{r}
model_rush <- aov(data = filter(data_taxa, year == "2015"),#, type == "meadow"), 
                  formula = rush ~ type + location + replicate + treat_plot) 

model_rush_int <- aov(data = filter(data_taxa, year == "2015"),#, type == "meadow"), 
                      formula = rush ~ type * location * replicate * treat_plot) 

summary(model_rush)

summary(model_rush_int)
```
There is significant difference between types (meadow and pasture).  So lets look at meadow and pasture separately: 

Meadow:
```{r}
model_rush_int <- aov(data = filter(data_taxa, year == "2015", type == "meadow"), 
                      formula = rush ~ location * replicate * treat_plot) 

summary(model_rush_int)
```
There is no significant difference in the main effects, but there is  in the interaction between location and treatment, and replicate and treatment.  

post hoc
```{r}
mead_tuk <- TukeyHSD(model_rush_int, "location:treat_plot")
mead_tuk <- data.frame(mead_tuk$`location:treat_plot`)
mead_tuk %>% arrange(p.adj)
```
only between one combination of 

####Pasture:
```{r}
model_rush_int <- aov(data = filter(data_taxa, year == "2015", type == "pasture"), 
                      formula = rush ~ location * replicate * treat_plot) 

summary(model_rush_int)
```
There is no significant difference in the main effects, but there is  in the interaction between location and treatment, and replicate and treatment.  

post hoc
```{r}
TukeyHSD(model_rush_int, "location:treat_plot")
```

###Looking at year as a factor

#meadow
```{r}
model_rush <- aov(data = filter(data_taxa, type == "meadow"), 
                  formula = rush ~ location + replicate + treat_plot + year) 

model_rush_int <- aov(data = filter(data_taxa, type == "meadow"), 
                      formula = rush ~ location * replicate * treat_plot * year) 

summary(model_rush)

summary(model_rush_int)
```


```{r}
model_rush <- aov(data = filter(data_taxa, type == "meadow"), 
                  formula = rush ~ treat_plot + year) 

model_rush_int <- aov(data = filter(data_taxa, type == "meadow"), 
                      formula = rush ~ treat_plot * year) 

summary(model_rush)

summary(model_rush_int)
```


post hoc
```{r}
TukeyHSD(model_rush, "treat_plot")
```

#pasture
```{r}
model_rush <- aov(data = filter(data_taxa, type == "pasture"), 
                  formula = rush ~ location + replicate + treat_plot + year) 

model_rush_int <- aov(data = filter(data_taxa, type == "pasture"), 
                      formula = rush ~ location * replicate * treat_plot * year) 

summary(model_rush)

summary(model_rush_int)
```


```{r}
model_rush <- aov(data = filter(data_taxa, type == "pasture"), 
                  formula = rush ~ treat_plot + year) 

model_rush_int <- aov(data = filter(data_taxa, type == "pasture"), 
                      formula = rush ~ treat_plot * year) 

summary(model_rush)

summary(model_rush_int)
```
No sign diffs at all. 



##Nested design

`replicate` is nested within `type`.  `treat-plot` is not? nested because ??? .  Hang on, is treat-plot actually two variables - or is replicate a variable at all ???? aaaagh

###TWo-level nested design ANOVA

`/` means a factor is nested in another. (Dytham p197)
```{r}

model_rush <- aov(data = filter(data_taxa, year == "2015", type == "meadow"), 
                  formula = rush ~ location/replicate + treat_plot) 


summary(model_rush)
```



```{r}

model_rush <- aov(data = filter(data_taxa, type == "meadow"), 
                  formula = rush ~ location/replicate + treat_plot + year) 


summary(model_rush)
```


###repeated measures ANOVA 

But error needs to be specified differently because `year` is a repeated measure (Dytham p128)
```{r}

model_rush <- aov(data = filter(data_taxa, type == "meadow"), 
                  formula = rush ~ treat_plot + location * year + Error(rush/((treat_plot + location) * year))) 


summary(model_rush)
```

(Mangiafico, S)[https://rcompanion.org/rcompanion/d_07.html]



```{r}
summary(aov(data = filter(data_taxa, type == "meadow"), 
    formula = rush ~ treat_plot))

```



```{r}
summary(aov(data = filter(data_taxa, type == "meadow", ), 
    formula = rush ~ treat_plot))

```
























