---
title: "Rush Trial: baseline similarity indices"
output: html_notebook
---

#load packages
```{r}
library(tidyverse)
library(vegan)
library(vegan3d)
library(ecodist)
```

#load data

```{r}
spp_obs <- read.csv("../data/prepped/spp_obs_w.csv", header = TRUE)
spp_obs$X <- NULL
```

#prepare data
filter out just the baseline year 
```{r}
spp_obs <- filter(spp_obs, year == "2015")
```

split observations and site data
```{r}
species <- spp_obs[8:ncol(spp_obs)]
site <- spp_obs[1:7]
```

remove empty species
```{r}
species <- species[,-which(colSums(species) == 0)]
```

#Calculate similarity

##Create distance matrices
```{r}
dist_all<- vegdist(species, method = "bray", binary = TRUE) #use cover values: binary = F 
                                                            #use pres/abs: binary = T
hist(dist_all)
boxplot(dist_all)
```


hang on, the dist_all matrix includes all comparisons, so I can turn it into a matrix object, then subset...
```{r}
dist_mat <- as.matrix(dist_all)

dist_mead <- as.dist(dist_mat[
  which(site$type == "meadow"),
  which(site$type == "meadow")
])
range(dist_mead)

dist_past <- as.dist(dist_mat[
  which(site$type == "pasture"),
  which(site$type == "pasture")
])


dist_HP <- as.dist(dist_mat[
  which(site$location == "HP"),
  which(site$location == "HP")
])

dist_VP <- as.dist(dist_mat[
  which(site$location == "VP"),
  which(site$location == "VP")
])


dist_HP <- as.dist(dist_mat[
  which(site$location == "HP"),
  which(site$location == "HP")
])

dist_HM <- as.dist(dist_mat[
  which(site$location == "HM"),
  which(site$location == "HM")
])


dist_HP <- as.dist(dist_mat[
  which(site$location == "HP"),
  which(site$location == "HP")
])

dist_LM <- as.dist(dist_mat[
  which(site$location == "LM"),
  which(site$location == "LM")
])


```

make a dataframe with distance results for comparison
```{r}
dist_df <- cbind(dist = c(dist_all, dist_mead, dist_past), 
                 comparison = c(rep("all", length(dist_all)),
                                rep("meadow", length(dist_mead)), 
                                rep("pasture", length(dist_past))))
dist_df <- as.data.frame(dist_df)
dist_df$dist <- as.character(dist_df$dist)
dist_df$dist <- as.numeric(dist_df$dist)
```

plot it
```{r}
boxplot(dist_df$dist ~ dist_df$comparison)
boxplot(dist_df$dist[which(dist_df$comparison == "all")])
```

do it again for locations
```{r}
dist_df_mead <- cbind(dist = c(dist_mead, dist_HM, dist_LM), 
                 comparison = c(rep("all_mead", length(dist_all)),
                                rep("HM", length(dist_HM)), 
                                rep("LM", length(dist_LM))))
dist_df_mead <- as.data.frame(dist_df_mead)
dist_df_mead$dist <- as.character(dist_df_mead$dist)
dist_df_mead$dist <- as.numeric(dist_df_mead$dist)

boxplot(dist_df_mead$dist ~ dist_df_mead$comparison)
```

```{r}
dist_df_past <- cbind(dist = c(dist_past, dist_HP, dist_VP), 
                 comparison = c(rep("all_past", length(dist_past)),
                                rep("HP", length(dist_HP)), 
                                rep("VP", length(dist_VP))))
dist_df_past <- as.data.frame(dist_df_past)
dist_df_past$dist <- as.character(dist_df_past$dist)
dist_df_past$dist <- as.numeric(dist_df_past$dist)

boxplot(dist_df_past$dist ~ dist_df_past$comparison)
```

```{r}
dist_past_HP <- sapply(seq(1:8), function(tp) 
  as.dist(dist_mat[
    which(site$location == "HP" & site$treat_plot == tp),
    which(site$location == "HP" & site$treat_plot == tp)
    ]))



boxplot(dist_past_HP)

dist_df_past_HP <- cbind(dist = c(dist_past, dist_HP, c(dist_past_HP)), 
                 comparison = c(rep("all_past", length(dist_past)),
                                rep("all_HP", length(dist_HP)), 
                                rep(1:8, each = nrow(dist_past_HP))
                                ))
dist_df_past_HP <- as.data.frame(dist_df_past_HP)
dist_df_past_HP$dist <- as.character(dist_df_past_HP$dist)
dist_df_past_HP$dist <- as.numeric(dist_df_past_HP$dist)

boxplot(dist_df_past_HP$dist ~ dist_df_past_HP$comparison)
```
```

