---
title: "Rush Trial: statistical analysis"
output: html_notebook
---

#Packages
```{r}
library(tidyverse)
library(nortest)
library(ggpubr)
```

#Import data
```{r}
cover_data <- read.csv("../data/cover_data.csv", header = TRUE)
```


reminder of the taxa
```{r}
unique(cover_data$taxon)
```



#Test assumptions

## Homogeneity of variances

###Fligner killeen test

The Fligner-Killeen test: **if the p-value is < 0.05 then reject hypothesis that variances are homogeneous**

```{r}
cover_data %>% 
  filter(taxon == "rushes") %>% 
  fligner.test(x = .$cover, g = .$treat_plot)
```

No evidence that the variances of *rushes* accross the treatments are not homogeneous (Fligner-Killeen:med chi-squared = 5.7217, df = 7, p-value = 0.5726).


```{r}
cover_data %>% 
  filter(taxon == "grasses") %>% 
  fligner.test(x = .$cover, g = .$treat_plot)
```

No evidence that the variances of *grasses* accross the treatments are not homogeneous (Fligner-Killeen:med chi-squared = 12.738, df = 7, p-value = 0.07876).


```{r}
cover_data %>% 
  filter(taxon == "herbs") %>% 
  fligner.test(x = .$cover, g = .$treat_plot)
```

No evidence that the variances of *herbs* accross the treatments are not homogeneous (Fligner-Killeen:med chi-squared = 5.5001, df = 7, p-value = 0.5992).


```{r}
cover_data %>% 
  filter(taxon == "bryophytes") %>% 
  fligner.test(x = .$cover, g = .$treat_plot)
```

No evidence that the variances of *bryophytes* accross the treatments are not homogeneous (Fligner-Killeen:med chi-squared = 1.6505, df = 7, p-value = 0.9766).


```{r}
cover_data %>% 
  filter(taxon == "sedges") %>% 
  fligner.test(x = .$cover, g = .$treat_plot)
```

No evidence that the variances of *sedges* accross the treatments are not homogeneous (Fligner-Killeen:med chi-squared = 7.3397, df = 7, p-value = 0.3944).


##Normality

plot data
```{r}
ggplot(cover_data, aes(x = cover)) +
  geom_histogram() +
  facet_grid(facets = ~taxon)
```

not looking particularly normal


###Anderson Darling test

Anderson Darling composite test for normality: **if the p-value is < 0.05 then reject hypothesis that the distribution is normal**


```{r}
ad.test(x = cover_data$cover[which(cover_data$taxon == "rushes")])
```
nope



```{r}
ad.test(x = cover_data$cover[which(cover_data$taxon == "grasses")])
```
nope


```{r}
ad.test(x = cover_data$cover[which(cover_data$taxon == "bryophytes")])
```
well, yes


```{r}
ad.test(x = cover_data$cover[which(cover_data$taxon == "sedges")])
```
no way


```{r}
ad.test(x = cover_data$cover[which(cover_data$taxon == "herbs")])
```
nah


So basically, violates normality. 

oh, wait, lets transform: start again: 

plot data
```{r}
ggplot(cover_data, aes(x = log(cover))) +
  geom_histogram() +
  facet_grid(facets = ~taxon)
```

hard to say


###Anderson Darling test

Anderson Darling composite test for normality: **if the p-value is < 0.05 then reject hypothesis that the distribution is normal**


```{r}
ad.test(x = log(cover_data$cover[which(cover_data$taxon == "rushes")]))
```
nope



```{r}
ad.test(x = log(cover_data$cover[which(cover_data$taxon == "grasses")]))
```
nope


```{r}
ad.test(x = log(cover_data$cover[which(cover_data$taxon == "bryophytes")]))
```
worse


```{r}
ad.test(x = log(cover_data$cover[which(cover_data$taxon == "sedges")]))
```
no way


```{r}
ad.test(x = log(cover_data$cover[which(cover_data$taxon == "herbs")]))
```
noope



