---
title: "Rush Trial: explore baseline"
output: html_notebook
---
#load packages
```{r}
library(tidyverse)
library(vegan)
library(vegan3d)
```

#load data

```{r}
spp_obs <- read.csv("../data/prepped/spp_obs_w.csv", header = TRUE)
spp_obs$X <- NULL
```

#prepare data
filter out just the baseline year and only pasture

```{r}
spp_obs <- filter(spp_obs, year == "2015")
spp_obs <- filter(spp_obs, type == "pasture")
```

split observations and site data

```{r}
species <- spp_obs[8:ncol(spp_obs)]
site <- spp_obs[1:7]
```

remove empty species
```{r}
species <- species[,-which(colSums(species) == 0)]
```

maybe remove species occuring in only one quadrat
```{r}
species_g <- species %>% 
  gather(key = "species", value = "cover") %>% 
  filter(cover > 0)


species_g %>% 
  group_by(species) %>% 
  count() %>% 
  filter(n < 5)
```
hmm, not sure

how about maybe remove species that are under a particular cover threshold: 
```{r}
species_g <- species %>% 
  gather(key = "species", value = "cover") %>% 
  filter(cover > 0)


species_g %>% 
  group_by(species) %>% 
  summarise(cov_sum = sum(cover)) %>% 
  filter(cov_sum < 10)
```

not now

#clustering

```{r}
#create distance matrix
dist_spp <- vegdist(species, method = "bray")

#visualise

#single linkage clustering
csin <- hclust(dist_spp, method="single")
csin
plot(csin, hang=-1)

#complete linkage 
ccom <- hclust(dist_spp, method="complete")
ccom
plot(ccom, hang=-1)

#average linkage 
caver <- hclust(dist_spp, method="aver")
caver
plot(caver, hang=-1)
```


description of classes
community tables ordered by a dendrogram, ordination or environmental
variables
```{r}
#vegemite(species, caver)

#doesn't work because cover values are longer than 1.  suggested I use 'scale' but I don't know if I really need this
```

Number of classes
 change `n.class` to change number of classes
```{r}
n.class <- 10
plot(csin, hang=-1)
rect.hclust(csin, n.class)
plot(ccom, hang=-1)
rect.hclust(ccom, n.class)
plot(caver, hang=-1)
rect.hclust(caver, n.class)
```

extract cluster identities
```{r}
cl <- cutree(ccom, n.class)
cl
```

extract cluster identities
```{r}
cl <- cutree(ccom, n.class)

clusters <- site
clusters$cl.ccom <- cl

clusters %>% 
  arrange(cl.ccom)
```

Ordination and clustering

metric scaling a.k.a. principal coordinates analysis
(PCoA) that maps observed dissimilarities linearly onto low-dimensional graph
using the same dissimilarities we had in our clustering.

```{r}
ord <- cmdscale(dist_spp)
ordiplot(ord)
ordihull(ord, cl, lty=3)
ordispider(ord, cl, col="blue", label=TRUE)
ordiellipse(ord, cl, col="red")

```



Manipulating dendrogram

```{r}
den <- as.dendrogram(caver)
```

In the following we rearrange the dendrogram so that the ordering of leaves
corresponds as closely as possible with the first ordination axis:

```{r}
x <- scores(ord, display = "sites", choices = 1)
oden <- reorder(den, x)
#We plot the dendrorgrams together:
par(mfrow=c(2,1))
plot(den)
plot(oden)
par(mfrow=c(1,1))
```


The vegemite function has a graphical sister function tabasco2
that can also
display the dendrogram. Moreover, it defaults to rerrange the dendrogram by
the first axis of Correspondence Analysis:
```{r}
tabasco(x = species)#, caver)
```

```{r}
species[which(rowSums(species)<=0),]
```

#Ordination

```{r}
library(vegan)
library(MASS)
```

DECORANA
```{r}
dca.1 <- decorana(species) #carry out ordination
dca.1
```

Non-metric Multidimensional scaling
```{r}
mds.1 <- metaMDS(dist_spp)
```



```{r}
ordiplot(dca.1, display = "sites", type = "points")
ordiplot(mds0.1, display = "sites", type = "t")
```

# Constrained ordination

Canonical Correspondence Analysis
```{r}
cca.1 <- cca(species ~ location + treat_plot, site)
cca.1
```

```{r}
plot(cca.1, display = "sites")
```

3d
```{r}
ordiplot3d(cca.1,display = "sites", type = "h")
ordirgl(cca.1, cex = 0.5)
```


amazing 3d

```{r}
if (interactive() && require(rgl, quietly = TRUE)) {
  ord <- rda(decostand(species, "hellinger"))
  ordirgl(ord, size=4, col = "yellow")
  orgltext(ord, display = "sites")
  with(site, ordirgl(ord, col = as.numeric(treat_plot), scaling = "sites"))
  ## orglspider & orglellipse: col by groups
  with(site, orglspider(ord, location, col = 1:3, scaling = "sites"))
#  with(site, orglellipse(ord, location, col = 1:3, kind = "se", 
#                         conf = 0.95, scaling = "sites"))
}
```






