---
title: 'Rush Trial: analyse one by one:'
output:  #html_notebook
  html_document:
    keep_md: yes

---

##Packages
```{r}
library(tidyverse)
library(skimr)
library(funModeling)

library(ggpubr)   # helps make publication-ready graphs
library(broom)    # cleans up code output from common functions

library(nortest)
```

##Import data
```{r}
grp_calc_w <- read.csv("../data/prepped/taxon_grp_calc_w.csv", header = TRUE)
```

```{r}
#drop unwanted columns
data_taxa <- grp_calc_w %>% 
  select(-c(X, uid)) 

#make categories factors
data_taxa <- data_taxa %>% 
  mutate_at(c("year", "treat_plot"), as.factor)

data_taxa_long <- data_taxa %>% 
  select(-cover_tot) %>% 
  gather(key = "taxon", value = "cover", grass:woodrush) %>% 
  mutate(taxon, taxon = as.factor(taxon))

head(data_taxa_long)

```



###Test for normality

####Anderson Darling test
Anderson Darling composite test for normality: **if the p-value is < 0.05 then reject hypothesis that the distribution is normal**

```{r}
#run Anderson Darling test
ad <- data_taxa_long %>% 
  group_by(type, location, taxon) %>% 
  summarise_at(vars(cover), funs(
    "A" = ad.test(.)$statistic,
    "p.val" = ad.test(.)$p.value,
    "formula" = ad.test(.)$data.name, 
    "n" = length(.)
    )) %>% 
  arrange(desc(p.val))

write.csv(ad, file = "../outputs/adTest.csv")

pander(filter(ad, p.val >= 0.05), digits = 3)
```

####Shapiro-Wilk test
Only available between 3 and 5000 samples.  **if the p-value is < 0.05 then reject hypothesis that the distribution is normal**
```{r}
#run  Shapiro-Wilk test
#doesn't seem to cope well with under represented taxa, so filtered out. 
shapiro <- data_taxa_long %>% 
  filter(taxon == "rush" | 
           taxon == "grass" | 
           taxon == "herb" | 
           taxon == "sedge") %>% 
  group_by(type, location, treat_plot, taxon) %>% 
  summarise_at(vars(cover), funs(
    "W" = shapiro.test(.)$statistic,
    "p.val" = shapiro.test(.)$p.value,
    "formula" = shapiro.test(.)$data.name, 
    "n" = length(.)
    )) %>% 
  arrange(desc(p.val))

write.csv(shapiro, file = "../outputs/shapiroTest.csv")

pander(filter(shapiro, p.val >= 0.05), digits = 3)
```

###Test for homoscedascity
Bartlett Test of Homogeneity of Variances: if the p-value is smaller than 0.05 then data not suitable for ANOVA
```{r}
#run Bartlett test for homogeneity of variances 

bartlett <- data_taxa_long %>% 
  filter(taxon == "rush" | 
           taxon == "grass" | 
           taxon == "herb" | 
           taxon == "sedge") %>% 
  group_by(type, location, treat_plot, taxon) %>% 
  summarise_at(vars(cover), funs(
    "Ksq" = bartlett.test(. ~ year)$statistic,
    "p.val" = bartlett.test(. ~ year)$p.value,
    "formula" = bartlett.test(. ~ year)$data.name, 
    "n" = length(.)
    )) %>% 
  arrange(desc(p.val))

write.csv(bartlett, file = "../outputs/bartlettTest.csv")

pander(filter(bartlett, p.val >= 0.05), digits = 3)
```

## Compare means: parametric
```{r}
aov <- data_taxa_long %>% 
  group_by(type, location, treat_plot, taxon) %>% 
  summarise_at(vars(cover), funs(
    "p.val" = tidy(aov(. ~ year))$p.value[1], 
    "n" = length(.)
  )) 

write.csv(aov, file = "../outputs/aov.csv")

pander(head(aov), digits = 3)
```

The following treatment plots have shown significant change: 

```{r}
aov.sig <- aov %>% 
  filter(p.val < 0.05) %>% 
  arrange(type, treat_plot, taxon) 

pander(aov.sig, digits = 3)
```

##Graph it

```{r}
#plot data, with facets shaded if p < 0.05

plotdata <- data_taxa %>% 
  filter(type == "meadow") %>% 
#  select(-one_of(c("woodrush", "moss", "horsetail", "cover_tot"))) %>% 
  gather(key = "taxon", value = "cover", grass:woodrush) #%>% 
#  mutate(taxon = factor(taxon, levels = c("rush", "grass", "herb", "sedge", "moss", "horsetail", "woodrush")))

#Create a data frame with the faceting variables
# and some dummy data (that will be overwritten)
facets <- filter(aov, type == "meadow")
facets$year <- 1
facets$cover <- 1


meadow.sm <-   ggplot(plotdata, 
                      aes(x = year, y = cover)) +
  geom_boxplot(data = plotdata, aes(group = year)) +
  geom_smooth(method = "lm", se = FALSE)   +
  scale_x_discrete(labels = c("15", "16", "17")) +
  geom_rect(data = subset(facets, p.val < 0.05), aes(fill = p.val), 
            xmin = -Inf,xmax = Inf,
            ymin = -Inf, ymax = Inf, alpha = 0.3) +
  facet_grid(taxon ~ treat_plot:location, scales = "free_y") +
  labs(title = "cover over time: meadow: ANOVA")

meadow.sm
ggsave(meadow.sm, filename = "../outputs/meadow.sm.aov.png", width = 30, units = "cm", dpi = 600)
```

```{r}
#plot data, with facets shaded if p < 0.05

plotdata <- data_taxa %>% 
  filter(type == "pasture") %>% 
#  select(-one_of(c("woodrush", "moss", "horsetail", "cover_tot"))) %>% 
  gather(key = "taxon", value = "cover", grass:woodrush) #%>% 
#  mutate(taxon = factor(taxon, levels = c("rush", "grass", "herb", "sedge", "moss", "horsetail", "woodrush")))

#Create a data frame with the faceting variables
# and some dummy data (that will be overwritten)
facets <- filter(aov, type == "pasture")
facets$year <- 1
facets$cover <- 1


pasture.sm <-   ggplot(plotdata, 
                      aes(x = year, y = cover)) +
  geom_boxplot(data = plotdata, aes(group = year)) +
  geom_smooth(method = "lm", se = FALSE)   +
  scale_x_discrete(labels = c("15", "16", "17")) +
  geom_rect(data = subset(facets, p.val < 0.05), aes(fill = p.val), 
            xmin = -Inf,xmax = Inf,
            ymin = -Inf, ymax = Inf, alpha = 0.3) +
  facet_grid(taxon ~ treat_plot:location, scales = "free_y") +
  labs(title = "cover over time: pasture: ANOVA")

pasture.sm
ggsave(pasture.sm, filename = "../outputs/pasture.sm.aov.png", width = 30, units = "cm", dpi = 600)
```


## Compare means: non-parametric

### Kruskal-Wallis rank sum test

This test estimates the probability that all the samples were drawn from the same population, i.e. that the true mean is the same for each survey.  
```{r}
#run Kruskal Wallis test
krusk <- data_taxa_long %>% 
  group_by(type, location, treat_plot, taxon) %>% 
  summarise_at(vars(cover), funs(
    "chi-sq" = kruskal.test(. ~ year)$statistic,
    "df" = kruskal.test(. ~ year)$parameter,
    "p.val" = kruskal.test(. ~ year)$p.value,
    #"method" = kruskal.test(. ~ year)$method, 
    "formula" = kruskal.test(. ~ year)$data.name, 
    "n" = length(.)
  )) 

write.csv(krusk, file = "../outputs/kruskalTest.csv")

pander(head(krusk), digits = 3)
```
The null hypothesis is that the samples are drawn from the same population, so if p is smaller than 0.05 we must reject that hypothesis. 


The following treatment plots have shown significant change: 

```{r}
krusk.sig <- krusk %>% 
  filter(p.val < 0.05) %>% 
  arrange(type, treat_plot, taxon) 

pander(krusk.sig, digits = 3)
```

##Graph it

```{r}
#plot data, with facets shaded if p < 0.05

plotdata <- data_taxa %>% 
  filter(type == "meadow") %>% 
#  select(-one_of(c("woodrush", "moss", "horsetail", "cover_tot"))) %>% 
  gather(key = "taxon", value = "cover", grass:woodrush) #%>% 
#  mutate(taxon = factor(taxon, levels = c("rush", "grass", "herb", "sedge", "moss", "horsetail", "woodrush")))

#Create a data frame with the faceting variables
# and some dummy data (that will be overwritten)
facets <- filter(krusk, type == "meadow")
facets$year <- 1
facets$cover <- 1


meadow.sm <-   ggplot(plotdata, 
                      aes(x = year, y = cover)) +
  geom_boxplot(data = plotdata, aes(group = year)) +
  geom_smooth(method = "lm", se = FALSE)   +
  scale_x_discrete(labels = c("15", "16", "17")) +
  geom_rect(data = subset(facets, p.val < 0.05), aes(fill = p.val), 
            xmin = -Inf,xmax = Inf,
            ymin = -Inf, ymax = Inf, alpha = 0.3) +
  facet_grid(taxon ~ treat_plot:location, scales = "free_y") +
  labs(title = "cover over time: meadow: kruskal wallis")

meadow.sm
#ggsave(meadow.sm, filename = "../outputs/meadow.sm.krusk.png", width = 30, units = "cm", dpi = 600)
```

```{r}
#plot data, with facets shaded if p < 0.05

plotdata <- data_taxa %>% 
  filter(type == "pasture") %>% 
#  select(-one_of(c("woodrush", "moss", "horsetail", "cover_tot"))) %>% 
  gather(key = "taxon", value = "cover", grass:woodrush) #%>% 
#  mutate(taxon = factor(taxon, levels = c("rush", "grass", "herb", "sedge", "moss", "horsetail", "woodrush")))

#Create a data frame with the faceting variables
# and some dummy data (that will be overwritten)
facets <- filter(krusk, type == "pasture")
facets$year <- 1
facets$cover <- 1


pasture.sm <-   ggplot(plotdata, 
                      aes(x = year, y = cover)) +
  geom_boxplot(data = plotdata, aes(group = year)) +
  geom_smooth(method = "lm", se = FALSE)   +
  scale_x_discrete(labels = c("15", "16", "17")) +
  geom_rect(data = subset(facets, p.val < 0.05), aes(fill = p.val), 
            xmin = -Inf,xmax = Inf,
            ymin = -Inf, ymax = Inf, alpha = 0.3) +
  facet_grid(taxon ~ treat_plot:location, scales = "free_y") +
  labs(title = "cover over time: pasture: kruskal wallis")

pasture.sm
#ggsave(pasture.sm, filename = "../outputs/pasture.sm.krusk.png", width = 30, units = "cm", dpi = 600)
```

